---
title: "Reproducible Research: Peer Assessment 1"
author: "Marian Drumea"
date: "October 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "figure/")
```

## Loading and preprocessing the data
```{r}
mydata <- read.csv("activity.csv")
myCleanData <- mydata
myCleanData[is.na(mydata$steps), 1] <- 0
sums <- aggregate(myCleanData[ ,1], list(date = myCleanData$date), sum)
```

## What is mean total number of steps taken per day?
### 1. Histogram of the total number of steps taken each day
```{r}
barplot(sums$x, names.arg = paste(substr(sums$date, 9, 10), "\n", substr(sums$date, 6, 7)), main = "Number of steps per day", xlab = "Date (DD/MM, all dates in 2012)", ylab = "Number of steps", cex.names = 0.75)
```

### 2. Mean/median total number of steps taken per day
```{r}
mean(sums$x)
median(sums$x)
```

## What is the average daily activity pattern?
### 1. Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
sumsInterval <- aggregate(myCleanData[ ,1], list(interval = myCleanData$interval), mean)
plot(sumsInterval$interval, sumsInterval$x, type = "l", names(sumsInterval$interval), main = "Average Number of Steps Across All Days, by Interval", xlab = "Interval", ylab = "Number of Steps") #to finish
```

### 2. 5-minute interval containing the maximum number of steps - 8:35 - 8:40
```{r}
sumsInterval[sumsInterval$x == max(sumsInterval$x), 1]
```

## Imputing missing values
### 1. Total number of missing values in the dataset
```{r}
sum(is.na(mydata$steps))
```

### 2 - 3. Imputing strategy used - replacing missing values with the overall mean for the interval 
```{r}
for (index in which(is.na(mydata$steps)))  {
  mydata[index, 1] <- sumsInterval[sumsInterval$interval == mydata[index, 3], 2]
}
```

### 4.1. Histogram for imputed dataset 
```{r}
sumsImputed <- aggregate(mydata[ ,1], list(date = mydata$date), sum)
barplot(sumsImputed$x, names.arg = paste(substr(sumsImputed$date, 9, 10), "\n", substr(sumsImputed$date, 6, 7)), main = "Number of Steps per Day", xlab = "Date (DD/MM, all dates in 2012)", ylab = "Number of Steps", cex.names = 0.75)
```

### 4.2. Mean/median total number of steps taken per day for imputed dataset
```{r}
mean(sumsImputed$x)
median(sumsImputed$x)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
mydata$weekday <- sapply(mydata$date, function(x) ifelse(weekdays(as.Date(x), abbreviate = TRUE) %in% c("Sat", "Sun"), "weekend", "weekday"))
sumsWeekday <- aggregate(mydata[ ,1], list(interval = mydata$interval, weekday = mydata$weekday), mean)
par(mfrow=c(2,1))
plot(as.numeric(sumsWeekday[sumsWeekday$weekday == "weekday", ]$interval), sumsWeekday[sumsWeekday$weekday == "weekday", ]$x, type = "l", main = "Number of Steps per Day During Weekdays", xlab = "Interval", ylab = "Number of Steps")
plot(as.numeric(sumsWeekday[sumsWeekday$weekday != "weekday", ]$interval), sumsWeekday[sumsWeekday$weekday != "weekday", ]$x, type = "l", main = "Number of Steps per Day During Weekends", xlab = "Interval", ylab = "Number of Steps")
```